RationalGRL: A Framework for Argumentation and Goal Modeling

### 2017-07-03 background.tex
added a paragraph to intro

Atkinson et al.~\cite{atkinson2007} develop an argument scheme for practical reasoning, which they call the \emph{practical reasoning argument scheme} (PRAS). This argument scheme is as follows\footnote{The original argument schemes uses slightly different terminology: we have replaced \emph{value} with \emph{softgoal}. Our results do not depend on these adjustments, but they make the relation between PRAS and GRL more clear.}:

@@

The capital letters $G$, $A$, and $S$ are variables, which can be instantiated with concrete goals, actions, and softgoals. Once instantiated, we obtain an argument. We can for instance instantiate the argument scheme above with the elements of Figure~\ref{fig:example-small} as follows:
\begin{itemize}
\item[] We have goal \texttt{Static simulation},
\item[] Doing action \texttt{Keep same cars} realizes goal \texttt{Static simulation}, 
\item[] Which contribute positively to the softgoal \texttt{Simulate} 
\item[] \textit{Therefore} 
\item[] We should perform action \texttt{Keep same cars}.
\end{itemize}

@@

These critical questions can point to new arguments that might counter the original argument. Take, for example, critical question 1: if we find that  \texttt{Keep same cars} actually does not realize goal \texttt{Static Simulation}, we can form a counter-argument.

@@

The usefulness of argument schemes, and PRAS in particular, for the analysis of practical reasoning situations has been shown in different areas such as e-democracy~\cite{cartwright2009IS}, law~\cite{atkinson2005legal}, planning \cite{medellin2013planning} and choosing between safety critical actions \cite{tolchinsky2012deliberation}. In this article, we aim at capturing the stakeholder's discussions as formal argumentation based on PRAS to decide whether intentional elements and their relationships are shown in the resulting goal model. This give a rationalization to the elements of the goal model in terms of underlying arguments, and furthermore, it allows one to understand why certain other elements have been rejected.

Argumentation schemes and their associated critical questions are very well suited for modeling discussions about a goal model: as Murukannaiah et al.~\cite{murukannaiah2015} have shown, they can guide users in systematically deriving conclusions and making assumptions explicit. This can also be seen from the obvious similaries between PRAS (actions, goals, values) and GRL (tasks, goals, softgoals) in the example above.

However, there are also some differences between PRAS and GRL. Not all elements and relationships of GRL fit into PRAS. For instance, PRAS does not have a notion of ``resource'', and many of the relationships of GRL do not occur in PRAS. Furthermore, it is not directly clear whether the critical questions as proposed by Atkinson actually apply to GRL. Therefore, we develop our own set of argument schemes and critical questions in Section~\ref{sect:gmas} by analyzing transcripts of discussions about the traffic simulator. Before doing so, we give an overview of our framework in the next section.

>>>>

Atkinson et al.~\cite{atkinson2007} develop an argument scheme for practical reasoning, which they call the \emph{Practical Reasoning Argument Scheme} (PRAS). This argument scheme is as follows\footnote{The original argument schemes uses slightly different terminology: we have replaced \emph{value} with \emph{softgoal}. Our results do not depend on these adjustments, but they make the relation between PRAS and GRL more clear.}:

@@

The capital letters $G$, $A$, and $S$ are variables, which can be instantiated with concrete goals, actions, and softgoals. Once instantiated, we obtain an argument. We can for instance instantiate the argument scheme above with the elements of Figure~\ref{fig:example-small} as follows: %SG: Need to update the example.
\begin{itemize}
\item[] We have goal \texttt{Static simulation},
\item[] Doing action \texttt{Keep same cars} realizes goal \texttt{Static simulation}, 
\item[] Which contribute positively to the softgoal \texttt{Simulate} 
\item[] \textit{Therefore} 
\item[] We should perform action \texttt{Keep same cars}.
\end{itemize}

@@

These critical questions can point to new arguments that might counter the original argument. Take, for example, critical question 1: if we find that  \texttt{Keep same cars} actually does not realize goal \texttt{Static Simulation}, we can form a counter-argument. %SG:placeholder for the examples.

@@

The usefulness of argument schemes, and PRAS in particular, for the analysis of practical reasoning situations has been shown in different areas such as e-democracy~\cite{cartwright2009IS}, law~\cite{atkinson2005legal}, planning \cite{medellin2013planning} and choosing between safety critical actions \cite{tolchinsky2012deliberation}. In this article, we aim at capturing the stakeholder's discussions as formal argumentation based on PRAS to decide whether intentional elements and their relationships are shown in the resulting goal model. This approach provides a rationalization to the elements of the goal model in terms of underlying arguments, and furthermore, it helps understanding why certain other elements have been rejected.

Argumentation schemes and their associated critical questions are very well suited for modeling discussions about a goal model: as Murukannaiah et al.~\cite{murukannaiah2015} have shown, they can guide users in systematically deriving conclusions and making assumptions explicit. This can also be seen from the obvious similarities between PRAS (actions, goals, values) and GRL (tasks, goals, softgoals) in the example above.

However, there are also some differences between PRAS and GRL. Not all elements and relationships of GRL fit into PRAS. For instance, PRAS does not have a notion of ``resource'', and many of the relationships of GRL do not occur in PRAS. Furthermore, it is not directly clear whether the critical questions as proposed by Atkinson can actually apply to GRL. Therefore, we develop our own set of argument schemes and critical questions in Section~\ref{sect:gmas} by analyzing transcripts of discussions about the traffic simulator. Before doing so, we give an overview of our framework in the next section.

### 2017-07-03 introduction.tex
added a paragraph to intro

%SG: We need to write a bit about the framework in past and refer it here. 
To answer the research question above and solve some of the concerns mentioned in Subsection~\ref{sect:problem_statement}, we propose a framework called RationalGRL~\cite{}. The RationalGRL framework combines the Goal-oriented Requirements Language (GRL) with a technique from argumentation and discourse modeling called \emph{argument schemes}~\cite{walton-etal2004}. %SG: we need to add 1-2 sentences about our previous work and then add what we do here... 
We develop a metamodel specifying how our framework extends GRL, and we use this as the basis for a prototype web-based implementation~\footnote{The tool is available at <GITHUB LINK>}. %SG: Placeholder for the link. 
We develop an initial list of argument schemes and critical questions by analyzing a set of transcripts containing discussion about an information system. Our framework is fully extensible, meaning that argument schemes and critical questions can be added by users for specific problem domains. We develop a methodology for using RationalGRL, which consists developing goal models and posing arguments in an integrated way. This methodology is intended to be used by practitioners in the field. We formalize RationalGRL using propositional logic, and develop algorithms for adding goal elements and arguments. We illustrates the steps of our framework  and how the framework can be applied in real cases with a traffic simulator example.

>>>>

\todo{Sepideh}{Marc}{Read this paragraph and see if it makes sense.}
To answer the research question above and solve some of the concerns mentioned in Subsection~\ref{sect:problem_statement}, we propose a framework called RationalGRL. The RationalGRL framework combines the Goal-oriented Requirements Language (GRL) with a technique from argumentation and discourse modeling called \emph{argument schemes}~\cite{walton-etal2004}.

In past, we introduced our preliminary steps towards RationalGRL framework~\cite{bnaic2014,vanzee-etal:renext2015,vanZee-etal:er2016} with its tool-support~\cite{vanZee-etal:comma2016}.  In our previous work, argument diagrams are integrated with GRL models to allow better representation of stakeholder's arguments and discussions. In our current paper, we improve RationalGRL framework by adding argument schemes and critical questions to argumentation framework and provide traceability links between the argumentation framework and GRL.  For this, we modify and extend our metamodel to specify how our framework extends GRL, and we use this as the basis for a prototype web-based implementation~\footnote{The tool is available at <GITHUB LINK>}. %SG: Placeholder for the link. 
We develop an initial list of argument schemes and critical questions by analyzing a set of transcripts containing discussion about an information system. Our framework is fully extensible, meaning that argument schemes and critical questions can be added by users for specific problem domains. We develop a methodology for using RationalGRL, which consists developing goal models and posing arguments in an integrated way. This methodology is intended to be used by practitioners in the field. We formalize RationalGRL using propositional logic, and develop algorithms for adding goal elements and arguments. We illustrates the steps of our framework  and how the framework can be applied in real cases with a traffic simulator example.
